---
title: "GR5261 Final Project"
author: "Group"
date: "2018.12.17"
output:
  word_document: default
  pdf_document: default
---

```{r}
packages.used=c("dplyr", "moments","tstools","TSA","ggplot2","copula","quadprog")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}


```

```{r,warning=FALSE}
library(dplyr)
library(moments)
library(tstools)
library(TSA)
library(ggplot2)
library(copula)
library(quadprog)
```

# Data Preparation

```{r}
# Loaded raw data and calculate return of each assets
raw <- read.csv("../data/Assets.csv")
raw <- raw[,-12]
dim(raw) # 60 * 12
raw1 <- raw[-60,]
raw2 <- raw[-1,]
netret <- raw2[,-1]/raw1[,-1]-1
logret <- log(netret+1)
netret$Date <- raw$Date[-1]
logret$Date <- raw$Date[-1]


### read  asset data
df <- read.csv("../data/Assets.csv")
TenAssets <- df[,2:11]
### Assumes that the log return of the assets follows a normal distribution, so we need to find out the log return of the ten assets
ReturnTenAssets <- apply(log(TenAssets),2,diff)


# SP500ETF returns (as market portfolio)
sp <- read.csv("../data/Assets.csv")
sp <- sp$SP500ETF
sp.netret <- sp[-1]/sp[-60]-1
sp.logret <- log(sp.netret+1)

# Risk free asset
rf <- read.csv("../data/Risk Free Rate.csv")
rf <- rf$X4.week.Treasury.bill.secondary.market.rate...discount.basis
rf <- rf[-1:-6]
rf <- rf[-60]
rf <- as.numeric(paste(rf))/100
rf.netret <- (1+rf)^(1/12)-1
rf.logret <- log(rf.netret+1)
risk_free_rate <- mean(rf.logret)
```

# 2. Descriptive Statistics
## Mean, sd, skewness, kurtoisis, beta
```{r}
descriptive <- matrix(NA, nrow = 5, ncol = 10)
y <- matrix(NA, 59, 11)
for (i in 1:10){
  descriptive[1, i] <- mean(logret[,i])
  descriptive[2, i] <- sqrt(var(logret[,i]))
  descriptive[3, i] <- skewness(logret[,i])
  descriptive[4, i] <- kurtosis(logret[,i])
  premium <- sp.logret - rf.logret
  y[,i] <- logret[,i] - rf.logret
  descriptive[5, i] <- as.numeric(lm(y[,i]~premium-1)$coefficients)
}
descriptive <- data.frame(descriptive)
colnames(descriptive) <- names(logret[1:10])
rownames(descriptive) <- c("Mean", "Standard Deviation", "Skewness", "Kurtosis", "Beta")
```

## Plot prices and returns
```{r}
ts.data <- ts(logret[,-11],start = c(2013,1), frequency = 12)
ts.raw <- ts(raw[,-1],start = c(2013,1), frequency = 12)
tt <- init_tsplot_theme(use_box = T)
for ( i in 1:10) {
  tt$lty <- c(1,3)
  tsplot(price = ts.raw[,i], tsr = ts.data[,i],theme = tt)
}
```

## Do the same for S&P 500/ compare it with assets
```{r}
tt <- init_tsplot_theme(use_box = T)
tt$lty <- c(1,3)
tsplot(price = ts(sp,start = c(2013,1), frequency = 12), tsr = ts(sp.logret,start = c(2013,1), frequency = 12), theme = tt)
```

## Histogram, boxplots, qq-plots, for each returns; 
```{r}
for (i in 1:10) {
  hist(logret[,i], xlab = 'Returns', main = paste("Histogram of" , names(logret[i])))
}

for (i in 1:10) {
  boxplot(logret[,i],  main = paste("boxplot of" , names(logret[i])))
}

for (i in 1:10) {
  qqnorm(logret[,i],  main = paste("Q-Q Plot of" , names(logret[i])))
}
```

## Construct pairwise scatter plots
```{r}
pairs(logret[1:10], gap = 0, pch = ".")
```

## Compute sample covariance matrix of returns on assets
```{r}
covariance <- cov(logret[1:10])
covariance
```

# 3. Portfolio Theory
## Markowitz approach
```{r}
mean_vect <- colMeans(ReturnTenAssets)
cov_mat <- cov(ReturnTenAssets)
sd_vect <- sqrt(diag(cov_mat))
```
### With shortsell
```{r}
#### no constraints on weights
M = length(mean_vect)
library(quadprog)
Amat = cbind(rep(1,M),mean_vect)
# The minimum and maximum returns that could be obtained are as follows
min_return <- -max(abs(mean_vect))-0.02
max_return <- max(abs(mean_vect))+0.02
muP = seq(min_return,max_return,length=50)
sdP = rep(0,50)
weights = matrix(0,nrow=50,ncol=M)
for (i in 1:length(muP))
{
result =
solve.QP(Dmat=cov_mat,dvec=rep(0,M), Amat=Amat,
c(1,muP[i]), meq=2)
sdP[i] = sqrt(2*result$value)
weights[i,] = result$solution
}
plot(sdP,muP,type="l",xlim=c(min(sdP)-0.025,max(sdP)+0.025),ylim=c(min(muP)-0.01,max(muP)+0.01))

mufree = risk_free_rate
points(0,mufree,cex=3,col="blue",pch="*")
sharpe =(muP-mufree)/sdP
ind = (sharpe == max(sharpe)) # locates the tangency portfolio
# weights of the tangency portfolio
tangency_portfolio_withSS <- weights[ind,] 
lines(c(0,sdP[ind]),c(mufree,muP[ind]),col="red",lwd=3)
points(sdP[ind],muP[ind],col="blue",cex=3,pch="*")
ind2 = (sdP == min(sdP))
MVP_withSS <- weights[ind2,] 
points(sdP[ind2],muP[ind2],col="green",cex=3,pch="*")
ind3 = (muP > muP[ind2])
lines(sdP[ind3],muP[ind3],type="l",xlim=c(0,.25),
ylim=c(0,.3),col="cyan",lwd=3)

legend("topright",c("efficient frontier","efficient portfolios",
 "tangency portfolio","min var portfolio"),
 lty=c(1,1,NA,NA),
 lwd=c(3,3,1,1),
 pch=c("","","*","*"),col=c("cyan","red","blue","green")
 )
```

### Without shortsell
```{r}
#### constraints on weights: all of the w is larger and equal than 0
Amat = cbind(rep(1,M),mean_vect,diag(1,nrow=M),diag(-1,nrow=M))
### Because of the constraints, the minimum and maximum returns that could be obtained are as follows
min_return2 <- min(mean_vect)
max_return2 <- max(mean_vect)
muP = seq(min_return2,max_return2,length=50)
sdP = rep(0,50)
weights = matrix(0,nrow=50,ncol=M)
for (i in 1:length(muP))
{
result =
solve.QP(Dmat=cov_mat,dvec=rep(0,M), Amat=Amat,
c(1,muP[i],rep(0,M),rep(-1,M)), meq=2)
sdP[i] = sqrt(2*result$value)
weights[i,] = result$solution
}
plot(sdP,muP,type="l",xlim=c(min(sdP)-0.02,max(sdP)+0.02),ylim=c(min(muP)-0.02,max(muP)+0.02))
mufree = risk_free_rate
points(0,mufree,cex=3,col="blue",pch="*")
sharpe =(muP-mufree)/sdP
ind_con = (sharpe == max(sharpe)) # locates the tangency portfolio
tangency_portfolio_withoutSS <- weights[ind_con,] # weights of the tangency portfolio
lines(c(0,sdP[ind_con]),c(mufree,muP[ind_con]),col="red",lwd=3)
points(sdP[ind_con],muP[ind_con],col="blue",cex=3,pch="*")
ind2_con = (sdP == min(sdP))
MVP_withoutSS <- weights[ind2_con,]
points(sdP[ind2_con],muP[ind2_con],col="green",cex=3,pch="*")
ind3_con = (muP > muP[ind2_con])
lines(sdP[ind3_con],muP[ind3_con],type="l",xlim=c(0,.25),

      ylim=c(0,.3),col="cyan",lwd=3)
legend("topright",c("efficient frontier","efficient portfolios",
"tangency portfolio","min var portfolio"),
lty=c(1,1,NA,NA),
lwd=c(3,3,1,1),
pch=c("","","*","*"),col=c("cyan","red","blue","green"),
pt.cex=c(1,1,3,3)
)
```

```{r}
# without shortsale
# weights of minimum variance portfolio
MVP_withoutSS
# weights of tangency portfolio
tangency_portfolio_withoutSS

# with shortsale
# weights of minimum variance portfolio
MVP_withSS
# weights of tangency portfolio
tangency_portfolio_withSS
```

# 4. Asset Allocation
```{r}
# Calculate the return of two tangency portfolios
tanretws <- tangency_portfolio_withSS %*% mean_vect
tanretwos <- tangency_portfolio_withoutSS %*% mean_vect
weights_tan_ws <- (0.005 - risk_free_rate) / (tanretws - risk_free_rate)
weights_tan_wos <- (0.005 - risk_free_rate) / (tanretwos - risk_free_rate)
round(rep(weights_tan_ws,10) * tangency_portfolio_withSS, 4)
round(rep(weights_tan_wos,10) * tangency_portfolio_withoutSS, 4)
```

# 5. Principle Component Analysis
```{r}
round(cor(ReturnTenAssets),4)
```

```{r}
eig <- eigen(cor(ReturnTenAssets))
#eig

#### conduct PCA
pca <- princomp(ReturnTenAssets,cor=TRUE)

pca$loadings
summary(pca)
```

```{r}

### plot for PCA
pca1 <- pca$loadings[,1]
pca5 <- pca$loadings[,5]
pca10 <- pca$loadings[,10]

dfPCA <- cbind(ReturnTenAssets%*%pca1,ReturnTenAssets%*%pca5,ReturnTenAssets%*%pca10)

dfRTA <- as.data.frame(cbind(1:59,ReturnTenAssets,dfPCA))
ggplot(data=dfRTA)+
  geom_line(mapping = aes(x = V1, y = V12,colour="PC01"))+
  geom_line(mapping = aes(x = V1, y = V13,colour="PC05"))+
  geom_line(mapping = aes(x = V1, y = V14,colour="PC10"))+
  labs(title = "PCA", x = "Month Index", y = "Principle Component")
  

### interpret PCA (first several to get 95% covariance, interpret the first three principle components)



### choose a protfolio based on the first principle components
### names: market factor portfolio
pca1 <- pca$loadings[,1]
PCA_weight <- pca1/sum(pca1)


```

## Factor Model
```{r}
### Fama 3 Factor Model
#### Here we use factor data from French's data library
### We use percent log return
FactorReturn <- 100*ReturnTenAssets

FactorAnalysis <- read.csv("../data/Factor.csv")
### Response: percent log return minus risk-free rate
Response <- FactorReturn-FactorAnalysis$RF

### First factor
# RM is the value_weighted monthly percent return on all the stocks minus one-month Treasury bill rate
RMRF <- FactorAnalysis$Mkt.RF
### Second factor
SMB <- FactorAnalysis$SMB
### Third factor
HML <- FactorAnalysis$HML

Apple <- lm(Response[,1]~RMRF+SMB+HML)
Apple
summary(Apple)

Delta <- lm(Response[,2]~RMRF+SMB+HML)
Delta
summary(Delta)

Exxon <- lm(Response[,3]~RMRF+SMB+HML)
Exxon
summary(Exxon)

GE <- lm(Response[,4]~RMRF+SMB+HML)
GE
summary(GE)

Honda <- lm(Response[,5]~RMRF+SMB+HML)
Honda
summary(Honda)

USSteel <- lm(Response[,6]~RMRF+SMB+HML)
USSteel
summary(USSteel)

Verizon <- lm(Response[,7]~RMRF+SMB+HML)
Verizon
summary(Verizon)

Walmart <- lm(Response[,8]~RMRF+SMB+HML)
Walmart
summary(Walmart)

CEA <- lm(Response[,9]~RMRF+SMB+HML)
CEA
summary(CEA)

Infosys <- lm(Response[,10]~RMRF+SMB+HML)
Infosys
summary(Infosys)

```

```{r}
round(PCA_weight, 4)
```

# 6. Risk Management
```{r}
# Now we have seven portfolio in total.
PortfolioWeights <- matrix(NA, nrow = 7, ncol = 10)
PortfolioWeights[1,] <- round(MVP_withSS,4)
PortfolioWeights[2,] <- round(MVP_withoutSS,4)
PortfolioWeights[3,] <- round(tangency_portfolio_withSS,4)
PortfolioWeights[4,] <- round(tangency_portfolio_withoutSS,4)
PortfolioWeights[5,] <- round(PCA_weight,4)
PortfolioWeights[6,] <- round(rep(weights_tan_ws,10) * tangency_portfolio_withSS,4)
PortfolioWeights[7,] <- round(rep(weights_tan_wos,10) * tangency_portfolio_withoutSS,4)
colnames(PortfolioWeights) <- colnames(logret)[1:10]
rownames(PortfolioWeights) <- c("MVP with short", "MVP without short", "tangency with short", "tangency without short", "PCA", "Combination with short", "Combination without short")
weights_tan_ws
weights_tan_wos
```

```{r}
CalRisk <- function(weights){
  mu <- as.numeric(descriptive[1,]) %*% as.matrix(as.numeric(weights), nrow = 10)
  sd <- sqrt(t(as.matrix(as.numeric(weights), nrow = 10)) %*% covariance %*%      
               as.matrix(as.numeric(weights), nrow = 10))
  annualmu <- 12 * mu
  annualsd <- sqrt(12) * sd
  normoutput <- qnorm(0.05)
  St <- 100000
  VaR <- St*(1-exp(mu+normoutput*sd))
  phi <- pnorm(normoutput-sd)
  ES <- St*(1-exp(mu+sd^2/2)*phi/0.05)
  portfolio_actual_ret <- as.matrix(logret[,1:10]) %*% 
    as.matrix(as.numeric(weights), nrow = 10)
  NPVaR <- -100000 * (exp(sort(portfolio_actual_ret)[3])-1)
  NPES <- -100000 * (exp((sort(portfolio_actual_ret)[3]
                         +sort(portfolio_actual_ret)[2]
                         +sort(portfolio_actual_ret)[1])/3)-1)
  SharpeRatio <- (mu - risk_free_rate) / sd
  return(c(mu, annualmu, sd, annualsd, VaR, ES, NPVaR, NPES, SharpeRatio))
}
```

```{r}
RiskParameter <- matrix(NA, nrow = 7, ncol = 9)
RiskParameter[1,] <- CalRisk(MVP_withSS)
RiskParameter[2,] <- CalRisk(MVP_withoutSS)
RiskParameter[3,] <- CalRisk(tangency_portfolio_withSS)
RiskParameter[4,] <- CalRisk(tangency_portfolio_withoutSS)
RiskParameter[5,] <- CalRisk(PCA_weight)

colnames(RiskParameter) <- c("monthlymean_ret", "yearlymean_ret", "monthlysd", "yearlysd", "5% VaR", "5% Expected Shortfall", "5% NPVaR", "5% NPExpected Shortfall", "Sharpe Ratio")
rownames(RiskParameter) <- c("MVP with short", "MVP without short", "tangency with short", "tangency without short", "PCA", "Combination with short", "Combination without short")
```

```{r}
# For portfolio combination with risk-free asset, in the actual data, risk-free asset also have volatility
CalRisk02 <- function(weights, weight02){
  mu <- as.numeric(descriptive[1,]) %*% as.matrix(as.numeric(weights), nrow = 10) + (1-weight02) * risk_free_rate
  sd <- sqrt(t(as.matrix(as.numeric(weights), nrow = 10)) %*% covariance %*%      
               as.matrix(as.numeric(weights), nrow = 10) + (1-weight02) * (1-weight02) * var(rf.logret))
  annualmu <- 12 * mu
  annualsd <- sqrt(12) * sd
  normoutput <- qnorm(0.05)
  St <- 100000
  VaR <- St*(1-exp(mu+normoutput*sd))
  phi <- pnorm(normoutput-sd)
  ES <- St*(1-exp(mu+sd^2/2)*phi/0.05)
  portfolio_actual_ret <- as.matrix(logret[,1:10]) %*% 
    as.matrix(as.numeric(weights), nrow = 10) + rep((1-weight02),59) * rf.logret
  NPVaR <- -100000 * (exp(sort(portfolio_actual_ret)[3])-1)
  NPES <- -100000 * (exp((sort(portfolio_actual_ret)[3]
                         +sort(portfolio_actual_ret)[2]
                         +sort(portfolio_actual_ret)[1])/3)-1)
  SharpeRatio <- (mu - risk_free_rate) / sd
  return(c(mu, annualmu, sd, annualsd, VaR, ES, NPVaR, NPES, SharpeRatio))
}
```

```{r}
RiskParameter[6,] <- CalRisk02(rep(weights_tan_ws,10) * tangency_portfolio_withSS, weights_tan_ws)
RiskParameter[7,] <- CalRisk02(rep(weights_tan_wos,10) * tangency_portfolio_withoutSS, weights_tan_wos)
```

## Bootstrap
```{r}
CalVarES1 <- function(portfolio){
  mu <- mean(portfolio)
  std <- sd(portfolio)
  normoutput <- qnorm(0.05)
  St <- 100000
  var <- St*(1-exp(mu+normoutput*std))
  
  phi <- pnorm(normoutput-std)
  
  ES <- St*(1-exp(mu+std^2/2)*phi/0.05)
  ES
  return(c(var,ES))
}
## function to calculate sde and 95% CI for var and ES based on bootstrap
CB <- function(port){
  varlist <- rep(NA,250)
  ESlist <- rep(NA,250)
  for (B in 1:250){
    index <- sample(1:length(port),replace=T)
    newPort <- port[index,]
    varlist[B] <- CalVarES1(newPort)[1]
    ESlist[B] <- CalVarES1(newPort)[2] 
  }
  sdeVar <- sd(varlist)
  sdeES <- sd(ESlist)
  var <- CalVarES1(port)[1]
  ES <- CalVarES1(port)[2]
  CIVar <- c(2*var-quantile(varlist,0.95),2*var-quantile(varlist,0.05))
  CIES <- c(2*ES-quantile(ESlist,0.95),2*ES-quantile(ESlist,0.05))
  return(c(sdeVar = sdeVar,sdeES = sdeES,CIVar= CIVar,CIES = CIES))
  
}
```

```{r}
BSOutput <- matrix(NA, nrow = 7, ncol = 6)
BSOutput[1,] <- CB(as.matrix(logret[,1:10]) %*% as.matrix(MVP_withSS, nrow = 10))
BSOutput[2,] <- CB(as.matrix(logret[,1:10]) %*% as.matrix(MVP_withoutSS, nrow = 10))
BSOutput[3,] <- CB(as.matrix(logret[,1:10]) %*% as.matrix(tangency_portfolio_withSS, nrow = 10))

BSOutput[4,] <- CB(as.matrix(logret[,1:10]) %*% as.matrix(tangency_portfolio_withoutSS, nrow = 10))
BSOutput[5,] <- CB(as.matrix(logret[,1:10]) %*% as.matrix(PCA_weight, nrow = 10))
BSOutput[6,] <- CB(as.matrix(logret[,1:10]) %*% 
    as.matrix(as.numeric(rep(weights_tan_ws,10) * tangency_portfolio_withSS), nrow = 10) + rep((1-weights_tan_ws),59)* rf.logret)
BSOutput[7,] <- CB(as.matrix(logret[,1:10]) %*% 
    as.matrix(as.numeric(rep(weights_tan_wos,10) * tangency_portfolio_withoutSS), nrow = 10) + rep((1-weights_tan_wos),59) * rf.logret)
colnames(BSOutput) <- c("sdeVar", "sdeES", "CIVar.95%", "CIVar.5%", "CIES.95%", "CIES.5%")
rownames(BSOutput) <- c("MVP with short", "MVP without short", "tangency with short", "tangency without short", "PCA", "Combination with short", "Combination without short")
```

# 7. Copulas
```{r}
## different copulas
## normal, t, clayon, gumbel,frank, joe
## gumbel and joe: can't have negative dependence
## clayton: remove points with small values of U1 and U2
### psuedo maximal likelihood method
### transfrom data to empirical CDFs so the first step is a nonparametric method in pseudo-Maximum likelihood
edata <- cbind(rank(ReturnTenAssets[,1])/60,rank(ReturnTenAssets[,2])/60,rank(ReturnTenAssets[,3])/60,rank(ReturnTenAssets[,4])/60,rank(ReturnTenAssets[,5])/60,rank(ReturnTenAssets[,6])/60,rank(ReturnTenAssets[,7])/60,rank(ReturnTenAssets[,8])/60,rank(ReturnTenAssets[,9])/60,rank(ReturnTenAssets[,10])/60)
ncop <- normalCopula(dim=10,dispstr="un")
Cnorm <- fitCopula(data=edata,copula=ncop,method="mpl")
AIC(Cnorm)

### t copoula
tcop <- tCopula(dim=10,dispstr="un")
Ct <- fitCopula(data=edata,copula=tcop,method="mpl")
AIC(Ct)

### Gumbel copula
Glcop <- archmCopula(family="gumbel",dim=10,param=2)
Cgumbel <- fitCopula(data=edata,copula=Glcop,method="mpl")
AIC(Cgumbel)

### Joe copula
Jcop <- archmCopula(family="joe",dim=10,param=2)
Cj <- fitCopula(data=edata,copula=Jcop,method="mpl")
AIC(Cj)

### Clayton copula
Clcop <- archmCopula(family="clayton",dim=10,param=2)
Ccl <- fitCopula(data=edata,copula=Clcop,method="ml")
AIC(Ccl)

### Frank copula
fcop <- archmCopula(family="frank",dim=10,param=2)
Cf <- fitCopula(data=edata,copula=fcop,method="ml")
AIC(Cf)

## For Gumbel and Joe copula, the dependence should be larger and equal than 0. 
##However, we have negative dependence on the returns, and this function automatically replace these with 0, but actually we should not consider it as the copula for our real data.
## Choose Clayton copula.
a <-c(AIC(Cnorm),AIC(Ct),AIC(Cgumbel),AIC(Cj),AIC(Ccl),AIC(Cf))
names(a) <- c("Normal Copula","T Copula","Gumbel Copula","Joe Copula","Clayton Copula","Frank Copula")

a

```



